---
title: "L01 Review"
subtitle: "Data Science II (STAT 301-2)"
author: "YOUR NAME HERE"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

The first goal of this lab is to ensure that all relevant software for this course is properly installed, functional, and up-to-date. Next, this lab aims to test student understanding of the fundamental modeling concepts (prediction vs. inference, classification vs. regression, and mechanistic vs. empirically driven models). Students are also asked to perform small EDAs, which should be review to those who took STAT 301-1.

*As a reminder: Students must create an R project for each lab, do their work within the project, and upload their 3 files (.R, .Rmd, and .html) according to the course naming conventions as described on Canvas.*

## Exercises

### Exercise 1 

Download and install the most recent versions of R and RStudio. Install and load `tidyverse`, `tidymodels`, and `ISLR`.

```{r, message=FALSE}
# Load package(s) here!

```

### Exercise 2

Use the tidyverse `read_` function(s) to load the `Auto` and `College` datasets in the `/data/` subdirectory into R.

### Exercise 3

Perform a short EDA on the `Auto` dataset. 

*You can find the codebook for `Auto` by loading the `ISLR` package and running the command `?Auto`.*

### Exercise 4

Perform a short EDA on the `College` dataset.

*You can find the codebook for `College` by loading the `ISLR` package and running the command `?College`.*

### Exercise 5

Explain, in your own words, the difference(s) between a predictive model and an inferential model. When might you use one kind of model versus the other? Why?

### Exercise 6

For each of the following examples, please describe: (1) whether the scenario is a classification or regression problem; (2) whether the goal is inferential or predictive; (3) the sample size (n) and the number of predictors (p).

(a) We collect a set of data on the top 500 firms in the US. For each firm we record profit, number of employees, industry, and CEO salary. We are interested in understanding which factors affect CEO salary.

(b) We are considering launching a new product and wish to know whether it will be a success or a failure. We collect data on 20 similar products that were previously launched. For each product we have recorded whether it was a success or a failure, the price charged for the product, the marketing budget, the competition price, and ten other variables.

(c) We are interested in predicting the % change in the USD/Euro exchange rate in relation to the weekly changes in the world stock markets. We collect data once a week for all of 2012. For each week we record the % change in the USD/Euro rate, the % change in the US market, the % change in the British market, and the % change in the German market.

### Exercise 7

Provide two real-life examples -- one of a case where someone might use a mechanistic model, and one where they might use an empirically driven model. For each, explain why. These examples must differ from those in the book.

### Exercise 8

Come up with a classification problem that you could model with the `Auto` dataset. (You don't need to actually create a model.) Specify the outcome and suggest some predictor variables. What are n and p (sample size and number of predictors) for this model?

### Exercise 9

Come up with a regression problem that you could model with the `College` dataset. (You don't need to actually create a model.) Specify the outcome and suggest some predictor variables. What are n and p (sample size and number of predictors) for this model?